---
title: "Light logger"
author: "Areej Hassan"
course: "Thesis"
---


```{r load-packages, include=FALSE}
#libraries to be loaded
library(LightLogR)
#these packages are needed for the examples as shown below.
library(flextable)
library(dplyr)
library(ggplot2)
library(lubridate)
library(gtsummary)
```

#### Load the dataset

```{r}
path <- "D:/files/set1/"
files <- list.files(path, full.names = TRUE)
#show how many files are listes
length(files)

tz <- "Europe/Berlin"
pattern <- "^(\\d{3})"
data <- import$ActLumus(files, tz = tz, dst_adjustment = TRUE, auto.plot = FALSE, auto.id = pattern, remove_duplicates = TRUE)

```

#### Visualize the dataset

```{r}
#create a new object with the erroneous data removed
data2 <- data %>% filter_Date(start = "2024-01-01")
data2 %>% gg_overview()

```

#### fill gap

```{r}
#looking at the various measurement intervals. Most of the data is in 30 second intervals, but lots have 60s differences. Which is the base interval? Could the various files have started 30 seconds off to one another?
data2 %>% count_difftime()
data2 %>% dominant_epoch()

#because we would have many gaps with 30s intervals, we will aggregate the data to 1 minute intervals
data1 <- data2 %>% aggregate_Datetime(unit = "1 min")

#we can see that this leaves us with "only" 10 gaps, as that is the highest Id number
data1 %>% gap_finder(gap.data = TRUE) %>% tail()

#looking at the overview for the unaggregated (data2) and aggregated data (data3), we see no discernible difference in the overview. Note that gg_overview does not plot gaps that interrupt the time sequence for only one instant.
data1 %>% gg_overview()

```

#### Extract only one day

```{r}
data1$date <- as.Date(data1$Datetime,tz=tz)
data1$month <- month(data1$date)
data_Id101 <- data1 %>% filter(Id == 106 & date == "2024-07-11")
data_Id101 %>% gg_day()

```



#### Time above threshold

```{r echo=TRUE, message=FALSE, warning=FALSE}
duration_above_threshold(
  Light.vector = data_Id101$MEDI,
  Time.vector = data_Id101$Datetime,
  threshold = 250,
  epoch = "dominant.epoch",
  na.rm = FALSE
)

```

#### Time above threshold for month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  data_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(duration_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE))
  
  print(paste("time above threshold for month", x))
  print(data_month)
}

```


#### Brightest 10 hours of the day

```{r echo=TRUE, message=FALSE, warning=FALSE}
brightest101 = bright_dark_period(
  Light.vector = data_Id101$MEDI,
  Time.vector = data_Id101$Datetime,
  as.df = TRUE,
  timespan = "10 hours",
  epoch = "dominant.epoch"
)
brightest101
Onset <- brightest101$brightest_10h_onset
Offset <- brightest101$brightest_10h_offset

data_Id101 %>% 
  gg_day(aes_col = Datetime >= Onset & Datetime <= Offset) +
  guides(color = "none")
```

#### Darkest 10 hours of the day

```{r echo=TRUE, message=FALSE, warning=FALSE}
M10_wrong <- 
bright_dark_period(
  Light.vector = data_Id101$MEDI,
  Time.vector = data_Id101$Datetime,
  as.df = TRUE,
  period = "darkest",
  timespan = "10 hours",
  epoch = "dominant.epoch"
)
M10_wrong
Onset <- M10_wrong$darkest_10h_onset
Offset <- M10_wrong$darkest_10h_offset

data_Id101 %>% 
  gg_day(aes_col = Datetime >= Onset & Datetime <= Offset) +
  guides(color = "none")
```

#### Darkest 10 hours of the day with loop

```{r echo=TRUE, message=FALSE, warning=FALSE}
M10 <- 
bright_dark_period(
  Light.vector = data_Id101$MEDI,
  Time.vector = data_Id101$Datetime,
  as.df = TRUE,
  period = "darkest",
  timespan = "10 hours",
  loop = TRUE
)
M10
Onset <- M10$darkest_10h_onset
Offset <- M10$darkest_10h_offset

data_Id101 %>% 
  gg_day(aes_col = Datetime >= Onset | Datetime <= Offset) +
  guides(color = "none")
```


#### barroso lighting metrics

```{r echo=TRUE, message=FALSE, warning=FALSE}
barroso_lighting_metrics <- data_Id101 %>%
  dplyr::reframe(barroso_lighting_metrics(MEDI, Datetime, as.df = TRUE))
barroso_lighting_metrics

```

#### frequency crossing

```{r echo=TRUE, message=FALSE, warning=FALSE}
freq_crossing <- data_Id101 %>%
  dplyr::reframe(frequency_crossing_threshold(MEDI, threshold = 250, as.df = TRUE))
freq_crossing

```

#### frequency crossing month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  freq_cross_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(frequency_crossing_threshold(MEDI, threshold = 250, as.df = TRUE))
  
  print(paste("Frequency of crossing lx 250 in month", x))
  print(freq_cross_month)
}

```


#### interdaily variability

```{r echo=TRUE, message=FALSE, warning=FALSE}
interdaily_var <- data1 %>%
  dplyr::reframe(intradaily_variability(MEDI, Datetime, as.df = TRUE))
interdaily_var

```

#### interdaily variability per month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  interdaily_var_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(intradaily_variability(MEDI, Datetime, as.df = TRUE))
  
  print(paste("interdaily variability in month", x))
  print(interdaily_var_month)
}

```

#### Non-visual circadian response

```{r echo=TRUE, message=FALSE, warning=FALSE}
nvRC <- data_Id101 %>%
  dplyr::mutate(
    nvRC = nvRC(MEDI, LIGHT, Datetime)
  )
data_Id101 %>% gg_day()
plot(nvRC$Datetime,log(nvRC$MEDI)/max(log(nvRC$MEDI)))
points(nvRC$Datetime, nvRC$nvRC, col='green', pch=12)

```


#### Period above threshold

```{r echo=TRUE, message=FALSE, warning=FALSE}
per_above_thr <- 
data_Id101 %>%
  dplyr::reframe("Period <250lx" = period_above_threshold(MEDI, Datetime, "above", threshold = 250))
per_above_thr
```


#### Period above threshold per month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  per_above_thr_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(period_above_threshold(MEDI, Datetime, "above", threshold = 250))
  
  print(paste("Period above threshold per month", x))
  print(per_above_thr_month)
}

```

#### Pulses above threshold

```{r echo=TRUE, message=FALSE, warning=FALSE}
pul_above_thr <- 
data_Id101 %>%
  dplyr::reframe(pulses_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE))
pul_above_thr
```


#### Pulses above threshold per month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  pulse_above_thr_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(pulses_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE))
  
  print(paste("Pulses above threshold per month", x))
  print(pulse_above_thr_month)
}
```

#### disparity index

```{r echo=TRUE, message=FALSE, warning=FALSE}
data_disparity_idex = data1 %>% 
    summarize(
    DI = disparity_index(
      MEDI
    )
  )
data_disparity_idex
```


#### disparity index per month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  data_disparity_idex_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(disparity_index(
      MEDI
    ))
  
  print(paste("Disparity index per month", x))
  print(data_disparity_idex_month)
}
```

#### exponential moving average

```{r echo=TRUE, message=FALSE, warning=FALSE}
data_EMA = data_Id101 %>% dplyr::mutate(MEDI.EMA = exponential_moving_average(MEDI, Datetime, decay = "10
                                                                              min"))
#data_EMA
data_EMA %>%
  ggplot2::ggplot(ggplot2::aes(x = Datetime)) +
  ggplot2::geom_line(ggplot2::aes(y = MEDI), colour = "black") +
  ggplot2::geom_line(ggplot2::aes(y = MEDI.EMA), colour = "red")
```

#### Interdaily Stability

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  IS_monthly = data1 %>% filter(month == x) %>%
  dplyr::reframe(interdaily_stability(MEDI, Datetime))
  
  print(paste("Pulses above threshold per month", x))
  print(IS_monthly)
}

```


#### TAT_250

```{r echo=TRUE, message=FALSE, warning=FALSE}
TAT250_Id101 = data_Id101 %>% 
  summarize(
    TAT_250 = duration_above_threshold(
      Light.vector = MEDI,
      Time.vector = Datetime,
      threshold = 250
    )
  )
TAT250_Id101
```

#### TAT_250 per month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  TAT250_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(duration_above_threshold(
      Light.vector = MEDI,
      Time.vector = Datetime,
      threshold = 250
    ))
  
  print(paste("Disparity index per month", x))
  print(TAT250_month)
}
```

#### Threshold for given duration

```{r echo=TRUE, message=FALSE, warning=FALSE}
TAT250_inv = data_Id101 %>% 
  dplyr::reframe(
                threshold_for_duration(MEDI, Datetime, duration = "10260 sec"))
TAT250_inv
```


#### Timing above threshold

```{r echo=TRUE, message=FALSE, warning=FALSE}
time_abv_thr = data_Id101 %>%
  dplyr::reframe(timing_above_threshold(MEDI, Datetime, "above", 250, as.df = TRUE))
time_abv_thr
```

#### Timing above threshold per month

```{r echo=TRUE, message=FALSE, warning=FALSE}
for (x in unique(data1$month)) {
  time_abv_thr_month = data1 %>% filter(month == x) %>%
  dplyr::reframe(timing_above_threshold(MEDI, Datetime, "above", 250, as.df = TRUE))
  
  print(paste("Timing above threshold per month", x))
  print(time_abv_thr_month)
}
```

#### TAT_250

```{r echo=TRUE, message=FALSE, warning=FALSE}
#create a new column in the data set with the weekday
data$wDay <- wday(data$Datetime, label = TRUE, week_start = 1)

#group the data and calculate the metrics
TAT_250_wday_Id101 <- 
data %>% 
  group_by(wDay, .add = TRUE) %>% 
  summarize(
    TAT_250 = duration_above_threshold(
      Light.vector = MEDI,
      Time.vector = Datetime,
      threshold = 250
    ), .groups = "drop_last"
  )

TAT_250_wday_Id101
```
